---
title: "Physiological parameter analysis- statistics and plots"
author: "Coral NRG NRK"
date: "11 12 2024"
output: html_document
---

This is an example of the statistical analysis and plots on the parameter "Host carbohydrates" (Carbs_Host.mg.cm2.).
This code was used repeteadly across all physiological parameters. 
If normality assumptions were met, mixed linear model was performed.
If assumptions were not met, PERMANOVA was performed. 


```{r, echo=FALSE}
library(utils)
library(pacman)
library(lme4)
library(lmerTest)
library(MuMIn)
library(dplyr)
library(stats)
library(mvabund)
library(lmerTest)
library(mvabund)
library (dplyr)
library(car)
library(rstatix)
library(ggplot2)
library(tidyr)
library(visreg)
library(rstatix)
library(PMCMRplus)
library(PMCMR)
library(viridis)
library(RColorBrewer)
library(Rmisc)
library(cowplot)
library(utils)
library(vegan)
library(readr)
library(predictmeans)
library(emmeans)
```

```{r}
rm(list=ls())

Physio=read.csv("Physiology_table_clean_results.csv")
head(Physio)
str(Physio)
colnames(Physio)

Physio$Date = as.Date(Physio$Date, format=  "%d/%m/%Y")
Physio$TP= as.factor(Physio$TP)
Physio$Carbs_Host.mg.cm2. <- as.numeric(as.character(Physio$Carbs_Host.mg.cm2.))
Physio$Colony=as.factor(Physio$Colony)
str(Physio)

```


```{r}
Physio_Sty= Physio %>% filter(Species=="Stylophora")%>% drop_na(Carbs_Host.mg.cm2.)

Physio_Poc= Physio %>% filter(Species=="Pocillopora")%>% drop_na(Carbs_Host.mg.cm2.) 



TPs_names= c("Feb21", "Jul21", "Aug21",
             "Jan22", "July22", "July22", "Aug22", "Aug22",
             "Jul23", "Jul23", "Aug23", "Aug23",
             "Jan24", "Jun24", "Sep24")


```



Stylophora:



Stats:
```{r}
#check homoscadecity
plot(Physio_Sty$Carbs_Host.mg.cm2.)  
#Null hypothesis for Levenes: There is no significant difference in the variance between groups 
# i.e. if p = <0.05 there is significant differences in the variation and therefore violates the assumption of ANOVA
#homogeneous variances: p = >0.05
    
leveneTest(Carbs_Host.mg.cm2.~ TP, Physio_Sty) 

leveneTest(Carbs_Host.mg.cm2.~ Colony, Physio_Sty)
# not all are homogeneous
```

```{r}
H_carbs_Sty <- ggplot() +
  geom_boxplot(data = Physio_Sty, aes(x = TP, y = Carbs_Host.mg.cm2.)) +
  geom_point(data = Physio_Sty, aes(x = TP, y = Carbs_Host.mg.cm2., color = Colony), position = position_jitterdodge(), size = 2) +
  scale_color_manual(values = c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")) +
  theme_classic(base_size = 16) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 0.7)) +
  ylab(expression(Carbohydrates~(mg~cm^{-2}))) +
  xlab(NULL) + 
  scale_x_discrete(labels = TPs_names)+
  scale_y_continuous(limits = c(0,0.5)) #adjust the limits by the values


H_carbs_Sty


ggsave("H_carbs_Sty.jpeg", width = 20, height = 10, units = "cm", dpi = 600)
```

Check for normally distributed data within groups for the fixed factors time point and treatment
# 3. Shapiro.test on raw data groups

```{r}
## Null hyp: the distribution of the data is not significantly different from a normal distribution i.e. p = > 0.05 the data are normal

TP1_Sty = subset(Physio_Sty, TP =="1")
shapiro.test(TP1_Sty$Carbs_Host.mg.cm2.) 

TP2_Sty = subset(Physio_Sty, TP =="2")
shapiro.test(TP2_Sty$Carbs_Host.mg.cm2.)

TP3_Sty = subset(Physio_Sty, TP =="3")
shapiro.test(TP3_Sty$Carbs_Host.mg.cm2.)

TP4_Sty = subset(Physio_Sty, TP =="4")
shapiro.test(TP4_Sty$Carbs_Host.mg.cm2.)

TP5_Sty = subset(Physio_Sty, TP =="5")
shapiro.test(TP5_Sty$Carbs_Host.mg.cm2.)

TP6_Sty = subset(Physio_Sty, TP =="6")
shapiro.test(TP6_Sty$Carbs_Host.mg.cm2.)

TP7_Sty = subset(Physio_Sty, TP =="7")
shapiro.test(TP7_Sty$Carbs_Host.mg.cm2.)

TP8_Sty = subset(Physio_Sty, TP =="8")
shapiro.test(TP8_Sty$Carbs_Host.mg.cm2.)

TP9_Sty = subset(Physio_Sty, TP =="9")
shapiro.test(TP9_Sty$Carbs_Host.mg.cm2.)

TP10_Sty = subset(Physio_Sty, TP =="10")
shapiro.test(TP10_Sty$Carbs_Host.mg.cm2.)

TP11_Sty = subset(Physio_Sty, TP =="11")
shapiro.test(TP11_Sty$Carbs_Host.mg.cm2.)

TP12_Sty = subset(Physio_Sty, TP =="12")
shapiro.test(TP12_Sty$Carbs_Host.mg.cm2.)

TP13_Sty = subset(Physio_Sty, TP =="13")
shapiro.test(TP13_Sty$Carbs_Host.mg.cm2.)

TP14_Sty = subset(Physio_Sty, TP =="14")
shapiro.test(TP14_Sty$Carbs_Host.mg.cm2.)

TP15_Sty = subset(Physio_Sty, TP =="15")
shapiro.test(TP15_Sty$Carbs_Host.mg.cm2.)

```

```{r}
#If ANOVA assumptions were met:
Carbs_Host.mg.cm2._lmer = lmer(Carbs_Host.mg.cm2.~ TP + (1|Colony), data = Physio_Sty, REML=FALSE)
summary(Carbs_Host.mg.cm2._lmer)
```

```{r}
anova(Carbs_Host.mg.cm2._lmer)
residplot(Carbs_Host.mg.cm2._lmer, group = "TP")
lsmeans(Carbs_Host.mg.cm2._lmer, pairwise~TP, adjust="tukey")
```



```{r}
#If model assumptions are not met:

permanova_result = permanova.lmer(Carbs_Host.mg.cm2._lmer, nperm = 999)
print(permanova_result)
```

```{r}
pairwise_permanova <- function(data, response, factor, random = NULL, nperm = 999) {
  levels <- unique(data[[factor]])
  n <- length(levels)
  results <- matrix(NA, n, n, dimnames = list(levels, levels))
  
  for (i in 1:(n-1)) {
    for (j in (i+1):n) {
      tryCatch({
        subset_data <- data[data[[factor]] %in% c(levels[i], levels[j]), ]
        
        if (is.null(random)) {
          model <- lm(as.formula(paste(response, "~", factor)), data = subset_data)
        } else {
          model <- lmer(as.formula(paste(response, "~", factor, "+", random)), data = subset_data, REML = FALSE)
        }
        
        perm_result <- permanova.lmer(model, nperm = nperm)
        
        cat("Comparing", levels[i], "and", levels[j], "\n")
        print(perm_result)
        
        p_value <- perm_result$Perm.p
        
        results[i, j] <- results[j, i] <- p_value
      }, error = function(e) {
        cat("Error in comparison of", levels[i], "and", levels[j], ":", conditionMessage(e), "\n")
      })
    }
  }
  
  return(results)
}
```

```{r}
post_hoc_results_random <- pairwise_permanova(Physio_Sty, "Carbs_Host.mg.cm2.", "TP", random = "(1|Colony)", nperm = 999)

# View the results
print(post_hoc_results_random)
```









Pocillopora:
Stats:
```{r}
#check homoscadecity
plot(Physio_Poc$Carbs_Host.mg.cm2.)  
#Null hypothesis for Levenes: There is no significant difference in the variance between groups 
# i.e. if p = <0.05 there is significant differences in the variation and therefore violates the assumption of ANOVA
#homogeneous variances: p = >0.05
    
leveneTest(Carbs_Host.mg.cm2.~ TP, Physio_Poc) 

leveneTest(Carbs_Host.mg.cm2.~ Colony, Physio_Poc)
# not all are homogeneous
```

```{r}
H_carbs_Poc <- ggplot() +
  geom_boxplot(data = Physio_Poc, aes(x = TP, y = Carbs_Host.mg.cm2.)) +
  geom_point(data = Physio_Poc, aes(x = TP, y = Carbs_Host.mg.cm2., color = Colony), position = position_jitterdodge(), size = 2) +
  scale_color_manual(values = c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", "#44AA99", "#999933", "#882255", "#661100", "#6699CC", "#888888")) +
  theme_classic(base_size = 16) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, vjust = 0.7)) +
  ylab(expression(Carbohydrates~(mg~cm^{-2}))) +
  xlab(NULL) + 
  scale_x_discrete(labels = TPs_names)+
  scale_y_continuous(limits = c(0,0.25))  #adjust the limits by the values #adjust the limits by the values


H_carbs_Poc


ggsave("H_carbs_Poc.jpeg", width = 20, height = 10, units = "cm", dpi = 600)
```

Check for normally distributed data within groups for the fixed factors time point and treatment
# 3. Shapiro.test on raw data groups

```{r}
## Null hyp: the distribution of the data is not significantly different from a normal distribution i.e. p = > 0.05 the data are normal

TP1_Poc = subset(Physio_Poc, TP =="1")
shapiro.test(TP1_Poc$Carbs_Host.mg.cm2.) 

TP2_Poc = subset(Physio_Poc, TP =="2")
shapiro.test(TP2_Poc$Carbs_Host.mg.cm2.)

TP3_Poc = subset(Physio_Poc, TP =="3")
shapiro.test(TP3_Poc$Carbs_Host.mg.cm2.)

TP4_Poc = subset(Physio_Poc, TP =="4")
shapiro.test(TP4_Poc$Carbs_Host.mg.cm2.)

TP5_Poc = subset(Physio_Poc, TP =="5")
shapiro.test(TP5_Poc$Carbs_Host.mg.cm2.)

TP6_Poc = subset(Physio_Poc, TP =="6")
shapiro.test(TP6_Poc$Carbs_Host.mg.cm2.)

TP7_Poc = subset(Physio_Poc, TP =="7")
shapiro.test(TP7_Poc$Carbs_Host.mg.cm2.)

TP8_Poc = subset(Physio_Poc, TP =="8")
shapiro.test(TP8_Poc$Carbs_Host.mg.cm2.)

TP9_Poc = subset(Physio_Poc, TP =="9")
shapiro.test(TP9_Poc$Carbs_Host.mg.cm2.)

TP10_Poc = subset(Physio_Poc, TP =="10")
shapiro.test(TP10_Poc$Carbs_Host.mg.cm2.)

TP11_Poc = subset(Physio_Poc, TP =="11")
shapiro.test(TP11_Poc$Carbs_Host.mg.cm2.)

TP12_Poc = subset(Physio_Poc, TP =="12")
shapiro.test(TP12_Poc$Carbs_Host.mg.cm2.)

TP13_Poc = subset(Physio_Poc, TP =="13")
shapiro.test(TP13_Poc$Carbs_Host.mg.cm2.)

TP14_Poc = subset(Physio_Poc, TP =="14")
shapiro.test(TP14_Poc$Carbs_Host.mg.cm2.)

TP15_Poc = subset(Physio_Poc, TP =="15")
shapiro.test(TP15_Poc$Carbs_Host.mg.cm2.)

```

```{r}

Carbs_Host.mg.cm2._lmer = lmer(Carbs_Host.mg.cm2.~ TP + (1|Colony), data = Physio_Poc, REML=FALSE)
summary(Carbs_Host.mg.cm2._lmer)
```

```{r}
#If ANOVA assumptions were met:
anova(Carbs_Host.mg.cm2._lmer)
residplot(Carbs_Host.mg.cm2._lmer, group = "TP")
lsmeans(Carbs_Host.mg.cm2._lmer, pairwise~TP, adjust="tukey")
```



```{r}
#If model assumptions are not met:

permanova_result = permanova.lmer(Carbs_Host.mg.cm2._lmer, nperm = 999)
print(permanova_result)
```

```{r}
pairwise_permanova <- function(data, response, factor, random = NULL, nperm = 999) {
  levels <- unique(data[[factor]])
  n <- length(levels)
  results <- matrix(NA, n, n, dimnames = list(levels, levels))
  
  for (i in 1:(n-1)) {
    for (j in (i+1):n) {
      tryCatch({
        subset_data <- data[data[[factor]] %in% c(levels[i], levels[j]), ]
        
        if (is.null(random)) {
          model <- lm(as.formula(paste(response, "~", factor)), data = subset_data)
        } else {
          model <- lmer(as.formula(paste(response, "~", factor, "+", random)), data = subset_data, REML = FALSE)
        }
        
        perm_result <- permanova.lmer(model, nperm = nperm)
        
        cat("Comparing", levels[i], "and", levels[j], "\n")
        print(perm_result)
        
        p_value <- perm_result$Perm.p
        
        results[i, j] <- results[j, i] <- p_value
      }, error = function(e) {
        cat("Error in comparison of", levels[i], "and", levels[j], ":", conditionMessage(e), "\n")
      })
    }
  }
  
  return(results)
}
```

```{r}
post_hoc_results_random <- pairwise_permanova(Physio_Poc, "Carbs_Host.mg.cm2.", "TP", random = "(1|Colony)", nperm = 999)

# View the results
print(post_hoc_results_random)
```






















